paper

### 2024-09-18 paper.qmd
.

@@

predictions_limrots <- limrots.result$FDR < 0.01
predictions_limrots <- factor(predictions_limrots, levels = c(TRUE, FALSE))
true_labels_limrots <- ifelse(row.names(limrots.result$data) %in% TP , TRUE, FALSE)
true_labels_limrots <- factor(true_labels_limrots, levels = c(TRUE, FALSE))
conf_matrix_limrots <- confusionMatrix(predictions_limrots, true_labels_limrots)

@@

```{r results = "hide" , message = FALSE, warning = FALSE}
# Load necessary packages
library(LimROTS)
library(parallel)
# Load the dataset
data("UPS1.Case1")
print(UPS1.Case1)
# Set random seed for reproducibility
set.seed(1234, kind = "default")
# Set metadata and formula for LimROTS analysis
meta.info <- c("Conc." , "inj")
B <- 1000  # Number of bootstrap samples
K <- dim(UPS1.Case1)[1]/4 # Set the value for K based on the data size
K <- floor(K) 
num_cores <- 10  # Number of cores for parallel processing
cluster <- makeCluster(num_cores)  # Create a parallel cluster
group.name <- "Conc."
formula.str <- "~ 0 + Conc.+ inj + Conc.*inj"  # Formula for group comparison + injections
# Run LimROTS analysis with trend and robust settings enabled
limrots.result <- LimROTS(data.exp = UPS1.Case1, 
                          B = B, K = K, meta.info = meta.info, 
                          cluster = cluster, group.name = group.name, 
                          formula.str = formula.str, trend = TRUE, robust = TRUE)
saveRDS(limrots.result, "../LimROTS paper/UPS1.Case1.LimROTS.rds")
```

@@

```{r results = "hide" , message = FALSE, warning = FALSE}
# Load necessary packages
library(LimROTS)
library(parallel)
# Load the dataset
data("UPS1.Case1")
print(UPS1.Case1)
# Set random seed for reproducibility
set.seed(1234, kind = "default")
# Set metadata and formula for LimROTS analysis
meta.info <- c("Conc." , "inj")
B <- 1000  # Number of bootstrap samples
K <- dim(UPS1.Case1)[1]/4 # Set the value for K based on the data size
K <- floor(K) 
num_cores <- 30  # Number of cores for parallel processing
cluster <- makeCluster(num_cores)  # Create a parallel cluster
group.name <- "Conc."
formula.str <- "~ 0 + Conc.+ inj + Conc.*inj"  # Formula for group comparison + injections
# Run LimROTS analysis with trend and robust settings enabled
rots.result <- LimROTS(data.exp = UPS1.Case1, 
                          B = B, K = K, meta.info = meta.info, 
                          cluster = cluster, group.name = group.name, 
                          formula.str = formula.str, trend = TRUE, robust = TRUE, n.ROTS = TRUE)
saveRDS(rots.result, "../LimROTS paper/UPS1.Case1.ROTS.rds")
```

@@

@@

```{r}
library(tidyverse, quietly = T, warn.conflicts = T)

alpha_thresholds <- seq(0, 0.2, by = 0.001)
significant_limROTS <- sapply(alpha_thresholds, function(alpha) sum(limrots.result$FDR < alpha))
significant_ROTS <- sapply(alpha_thresholds, function(alpha) sum(rots.result$FDR < alpha))
significant_limma <- sapply(alpha_thresholds, function(alpha) sum(limma.result$adj.P.Val < alpha))

plot_data <- data.frame(Alpha = alpha_thresholds, significant_limROTS = significant_limROTS,
                        significant_ROTS = significant_ROTS , significant_limma = significant_limma)

@@

# Plot using ggplot2
ggplot(plot_data.melt, aes(x = Alpha, y = FDR, color = Method)) +
  geom_line( size = 1) +
  theme_minimal() +
  labs(title = "Significant Proteins vs FDR",
       x = "FDR",
       y = "Number of Significant Proteins") +
   geom_vline(xintercept = 0.01, linetype="dotted", 
                color = "red", size=1) + 
    geom_vline(xintercept = 0.05, linetype="dotted", 
                color = "red", size=1)
```

@@

```{r results = "hide" , message = FALSE, warning = FALSE}
formula.str <- "~ Conc. + tool + fake.batch"
data("UPS1.Case4")
data <- assay(UPS1.Case4)
meta.info <- data.frame(colData(UPS1.Case4), check.names = FALSE, row.names = row.names(colData(UPS1.Case4)))
meta.info$sample.id <- NULL
ancova.result <- data.frame()
for(i in 1:nrow(data)){
  df.temp <- data[i,]
  data.temp <- data.frame(response = as.numeric(df.temp) , Conc. = meta.info$Conc.,
                     tool = meta.info$tool, fake.batch = meta.info$fake.batch )
  ancova_model <- lm(formula(paste("response", formula.str)), data = data.temp)
  coef_table <- anova(ancova_model)
  coef_table <- data.frame(coef_table)
  significant_groups <- coef_table[rownames(coef_table) == "Conc." , ]
  ancova.result <- rbind(ancova.result, data.frame(protein = row.names(data)[i], p.value =significant_groups[, 5] ))
}
ancova.result$FDR <- p.adjust(ancova.result$p.value, method = "fdr")

@@

# Set metadata and formula for LimROTS analysis
meta.info <- "mixing.ratio"
B <- 1000  # Number of bootstrap samples
K <- dim(BM21.Case0)[1]/2  # Set the value for K based on the data size
K <- floor(K) 
num_cores <- 10  # Number of cores for parallel processing
cluster <- makeCluster(num_cores)  # Create a parallel cluster
group.name <- "mixing.ratio"
formula.str <- "~ 0 + mixing.ratio"  # Formula for group comparison

@@

```{r}
length(which(limrots.result$FDR < 0.05))
summary(limrots.result$q_values)
plot(limrots.result$q_values)
hist(limrots.result$FDR)

limrots.result.df <- data.frame(proteins = row.names(limrots.result$data),
                                LimROTS.FC = limrots.result$corrected.logf,
                                FDR = limrots.result$FDR)
```

@@

# Set metadata and formula for LimROTS analysis
meta.info <- "mixing.ratio"
B <- 1000  # Number of bootstrap samples
K <- dim(BM21.Case0)[1]/2 # Set the value for K based on the data size
K <- floor(K) 
num_cores <- 10  # Number of cores for parallel processing
cluster <- makeCluster(num_cores)  # Create a parallel cluster
group.name <- "mixing.ratio"
formula.str <- "~ 0 + mixing.ratio"  # Formula for group comparison

# Run LimROTS analysis with trend and robust settings enabled
rots.result <- LimROTS(data.exp = BM21.Case0, 
                          B = B, K = K, meta.info = meta.info, 
                          cluster = cluster, group.name = group.name, 
                          formula.str = formula.str, trend = TRUE, robust = TRUE, n.ROTS = T)
```

>>>>

## Run t-test

library(SummarizedExperiment, quietly = T, warn.conflicts = F)
data("UPS1.Case0")
data <- assay(UPS1.Case0)
meta.info <- data.frame(colData(UPS1.Case0), check.names = FALSE, row.names = row.names(colData(UPS1.Case0)))
t.test.result <- data.frame()

for(i in 1:nrow(data)){
  p <- t.test(data[i,1:12], data[i,13:24])
  t.test.result <- rbind(t.test.result, data.frame(protein = row.names(data)[i] ,p.value = p$p.value))
}

t.test.result$fdr <- p.adjust(t.test.result$p.value)
saveRDS(t.test.result, "../LimROTS paper/UPS1.Case0.t.test.rds")

@@

predictions_limrots <- limrots.result$q_values$qvalues < 0.01
predictions_limrots <- factor(predictions_limrots, levels = c(TRUE, FALSE))
true_labels_limrots <- ifelse(row.names(limrots.result$data) %in% TP , TRUE, FALSE)
true_labels_limrots <- factor(true_labels_limrots, levels = c(TRUE, FALSE))
conf_matrix_limrots <- confusionMatrix(predictions_limrots, true_labels_limrots)

@@

```{r results = "hide" , message = FALSE, warning = FALSE}
# Load necessary packages
library(LimROTS)
library(parallel)
# Load the dataset
data("UPS1.Case1")
print(UPS1.Case1)
# Set random seed for reproducibility
set.seed(1234, kind = "default")
# Set metadata and formula for LimROTS analysis
meta.info <- c("Conc." , "inj")
B <- 1000  # Number of bootstrap samples
K <- dim(UPS1.Case1)[1]/4 # Set the value for K based on the data size
K <- floor(K) 
num_cores <- 50  # Number of cores for parallel processing
cluster <- makeCluster(num_cores)  # Create a parallel cluster
group.name <- "Conc."
formula.str <- "~ 0 + Conc.+ inj + Conc.*inj"  # Formula for group comparison + injections
# Run LimROTS analysis with trend and robust settings enabled
limrots.result <- LimROTS(data.exp = UPS1.Case1, 
                          B = B, K = K, meta.info = meta.info, 
                          cluster = cluster, group.name = group.name, 
                          formula.str = formula.str, trend = TRUE, robust = TRUE)
saveRDS(limrots.result, "../LimROTS paper/UPS1.Case1.LimROTS.rds")
```

@@

```{r results = "hide" , message = FALSE, warning = FALSE}
# Load necessary packages
library(LimROTS)
library(parallel)
# Load the dataset
data("UPS1.Case1")
print(UPS1.Case1)
# Set random seed for reproducibility
set.seed(1234, kind = "default")
# Set metadata and formula for LimROTS analysis
meta.info <- c("Conc." , "inj")
B <- 1000  # Number of bootstrap samples
K <- dim(UPS1.Case1)[1]/4 # Set the value for K based on the data size
K <- floor(K) 
num_cores <- 50  # Number of cores for parallel processing
cluster <- makeCluster(num_cores)  # Create a parallel cluster
group.name <- "Conc."
formula.str <- "~ 0 + Conc.+ inj + Conc.*inj"  # Formula for group comparison + injections
# Run LimROTS analysis with trend and robust settings enabled
rots.result <- LimROTS(data.exp = UPS1.Case1, 
                          B = B, K = K, meta.info = meta.info, 
                          cluster = cluster, group.name = group.name, 
                          formula.str = formula.str, trend = TRUE, robust = TRUE, n.ROTS = TRUE)
saveRDS(rots.result, "../LimROTS paper/UPS1.Case1.ROTS.rds")
```

@@

formula.str <- "~Conc.+ inj + Conc.*inj"
data("UPS1.Case1")
data <- assay(UPS1.Case1)
meta.info <- data.frame(colData(UPS1.Case1), check.names = FALSE, row.names = row.names(colData(UPS1.Case1)))
meta.info$sample.id <- NULL
ancova.result <- data.frame()
for(i in 1:nrow(data)){
  df.temp <- data[i,]
  data.temp <- data.frame(response = as.numeric(df.temp) , Conc. = meta.info$Conc.,
                     inj = meta.info$inj)
  ancova_model <- lm(formula(paste("response", formula.str)), data = data.temp)
  coef_table <- anova(ancova_model)
  coef_table <- data.frame(coef_table)
  significant_groups <- coef_table[rownames(coef_table) == "Conc." , ]
  ancova.result <- rbind(ancova.result, data.frame(protein = row.names(data)[i], p.value =significant_groups[, 5] ))
}
ancova.result$FDR <- p.adjust(ancova.result$p.value, method = "fdr")
saveRDS(ancova.result, "../LimROTS paper/UPS1.Case1.ancova.rds")
```

@@

```{r}
library(tidyverse, quietly = T, warn.conflicts = T)
library(ggsci)

limrots.result <- readRDS("../LimROTS paper/UPS1.Case1.LimROTS.rds")
rots.result <- readRDS("../LimROTS paper/UPS1.Case1.ROTS.rds")
limma.result <- readRDS("../LimROTS paper/UPS1.Case1.limma.rds")
ancova.result <- readRDS("../LimROTS paper/UPS1.Case1.ancova.rds")

alpha_thresholds <- seq(0, 0.2, by = 0.001)
significant_limROTS <- sapply(alpha_thresholds, function(alpha) sum(limrots.result$q_values$qvalues < alpha))
significant_ROTS <- sapply(alpha_thresholds, function(alpha) sum(rots.result$FDR < alpha))
significant_limma <- sapply(alpha_thresholds, function(alpha) sum(limma.result$adj.P.Val < alpha))
significant_ancova <- sapply(alpha_thresholds, function(alpha) sum(ancova.result$FDR < alpha))

plot_data <- data.frame(Alpha = alpha_thresholds, significant_limROTS = significant_limROTS,
                        significant_ROTS = significant_ROTS , significant_limma = significant_limma  ,significant_ancova = significant_ancova)

@@

# Plot using ggplot2
ggplot(plot_data.melt, aes(x = Alpha, y = FDR, color = Method)) +
  geom_line( size = 1) +
  theme_minimal() +
  labs(title = "Significant Proteins vs FDR Case study 1",
       x = "FDR",
       y = "Number of Significant Proteins") +
   geom_vline(xintercept = 0.01, linetype="dotted", 
                color = "red", size=1) + 
    geom_vline(xintercept = 0.05, linetype="dotted", 
                color = "red", size=1)+
    scale_color_npg() + 
    theme(
    plot.title = element_text(face = "bold", color = "black"),
    axis.title = element_text(face = "bold", color = "black"),
    axis.text = element_text(face = "bold", color = "black"),
    axis.ticks = element_line(color = "black")
  )
```

@@

```{r results = "hide" , message = FALSE, warning = FALSE}
formula.str <- "~ Conc. + tool + fake.batch"
data("UPS1.Case4")
data <- assay(UPS1.Case4)
meta.info <- data.frame(colData(UPS1.Case4), check.names = FALSE, row.names = row.names(colData(UPS1.Case4)))
meta.info$sample.id <- NULL
ancova.result <- data.frame()
for(i in 1:nrow(data)){
  df.temp <- data[i,]
  data.temp <- data.frame(response = as.numeric(df.temp) , Conc. = meta.info$Conc.,
                     tool = meta.info$tool, fake.batch = meta.info$fake.batch )
  ancova_model <- lm(formula(paste("response", formula.str)), data = data.temp)
  coef_table <- anova(ancova_model)
  coef_table <- data.frame(coef_table)
  significant_groups <- coef_table[rownames(coef_table) == "Conc." , ]
  ancova.result <- rbind(ancova.result, data.frame(protein = row.names(data)[i], p.value =significant_groups[, 5] ))
}
ancova.result$FDR <- p.adjust(ancova.result$p.value, method = "fdr")
ancova.result <- saveRDS("../LimROTS paper/UPS1.Case2.ancova.rds")

@@

# Set metadata and formula for LimROTS analysis
meta.info <- "mixing.ratio"
B <- 1000  # Number of bootstrap samples
K <- dim(BM21.Case0)[1]/2  # Set the value for K based on the data size
K <- floor(K) 
num_cores <- 30   # Number of cores for parallel processing
cluster <- makeCluster(num_cores)  # Create a parallel cluster
group.name <- "mixing.ratio"
formula.str <- "~ 0 + mixing.ratio"  # Formula for group comparison

trace(LimROTS::LimROTS , edit = T)

trace(LimROTS::calculateFDR , edit = T)

@@

```{r}
length(which(limrots.result$FDR < 0.01))

length(which(limrots.result$BH.pvalue < 0.01))

summary(limrots.result$q_values)
plot(limrots.result$q_values)
hist(limrots.result$FDR)

limrots.result.df <- data.frame(row.names  = row.names(limrots.result$data),
                                LimROTS.FC = limrots.result$corrected.logf,
                                FDR = limrots.result$FDR, p.value = limrots.result$pvalue, bh = limrots.result$BH.pvalue)
```

@@

# Set metadata and formula for LimROTS analysis
meta.info <- "mixing.ratio"
B <- 1000  # Number of bootstrap samples
K <- dim(BM21.Case0)[1]/2 # Set the value for K based on the data size
K <- floor(K) 
num_cores <- 30  # Number of cores for parallel processing
cluster <- makeCluster(num_cores)  # Create a parallel cluster
group.name <- "mixing.ratio"
formula.str <- "~ 0 + mixing.ratio"  # Formula for group comparison

trace(LimROTS::calculateFDR , edit = T)

#trace(LimROTS , edit = T)
# Run LimROTS analysis with trend and robust settings enabled
rots.result <- LimROTS(data.exp = BM21.Case0, 
                          B = B, K = K, meta.info = meta.info, 
                          cluster = cluster, group.name = group.name, 
                          formula.str = formula.str, trend = TRUE, robust = TRUE, n.ROTS = T)
```
